---
title: "Data-Analysis-project"
author: "Peter Riccio"
date: "2024-11-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Analysis Project

```{r}
library(tidyverse)
library(dplyr)
library(janitor)
library(lubridate)
#install packages 
```

```{r}

md_pres_vice_data <- read_csv("data/md_pres_vice_data.csv") |> clean_names ()
pres_election_by_md_counties <-read_csv("data/pres_election_by_md_counties.csv") |> clean_names()
senate_by_county <- read_csv("data/senate_by_county.csv") |> clean_names ()
senate_overall <-read_csv("data/senate_overall_data_v3.csv") |> clean_names()
eligible_active_voters_by_county <- read_csv("data/Eligible_Active_Voters_by_County.csv")|> clean_names()

voter_registration <- read.csv("data/voter_registration.csv")|> clean_names()

elect_2020 <- read.csv("data/voter_data - 2020pres.csv")|> clean_names()
elect_2016 <- read.csv("data/voter_data - 2016pres.csv")|> clean_names()
elect_2012 <- read.csv("data/voter_data - 2012pres.csv")|> clean_names()
elect_2008 <- read.csv("data/voter_data - 2008pres.csv")|> clean_names()

```

**MD Pres/ Vice Data:**

+-------------------+----------------------------------------------------------+----------------------------------+--------------------------------------------+----------------------------------------------+-----------------------------------------------------+----------------------------------+-------------------------------------------------------+
| name              | party                                                    | early_voting                     | election_day                               | mail_in_ballot                               | provisional                                         | total                            | percentage                                            |
+-------------------+----------------------------------------------------------+----------------------------------+--------------------------------------------+----------------------------------------------+-----------------------------------------------------+----------------------------------+-------------------------------------------------------+
| Name of candidate | Democratic/ Republican? libertarian/ Green/ Unaffiliated | Number of people who voted early | Number of people who voted on election day | Number of people who mailed in their ballots | Number of people who filled out provisional ballots | Total Number of people who voted | The percentage of people who voted for each candidate |
+-------------------+----------------------------------------------------------+----------------------------------+--------------------------------------------+----------------------------------------------+-----------------------------------------------------+----------------------------------+-------------------------------------------------------+

\

**Presidential Election by MD Counties:**

+---------------------------+-----------------------------------------------------------+----------------------------------------------------------+---------------------------------------------------------------+------------------------------------------------------------------------+---------------------------------------------------------------------+
| Jurisdiction              | Kamala_d_harris_and_tim_walz_democratic                   | donald_j_trump_and_jd_vance_republican                   | Chase_oliver_and_mile_ter_maat_libertatiran                   | jill_ellen_stein_and_rudolph_ware_green                                | Robert_f_kennendy_jr_and_nicole_shanahan_unaffiliated               |
+---------------------------+-----------------------------------------------------------+----------------------------------------------------------+---------------------------------------------------------------+------------------------------------------------------------------------+---------------------------------------------------------------------+
| The names of the counties | Number of people who voted for Kamala Harris and Tim Walz | Number of people who voted for Donald Trump and JD Vance | Number of people who voted for Chase Oliver and Mike Ter-maat | Number of people who voted for Jill Ellen Stein and Rudolph Ware Green | Number of people who voted for Robert F Kennedy and Nicole Shanahan |
+---------------------------+-----------------------------------------------------------+----------------------------------------------------------+---------------------------------------------------------------+------------------------------------------------------------------------+---------------------------------------------------------------------+

**Senate Overall:**

+----------------------+--------------------------------------------------+--------------------------------------------+-------------------------------------------+
| Jurisdiction         | angela_alsobrooks_democratic                     | larry_hogan_republican                     | Mike_scott_libertarian                    |
+----------------------+--------------------------------------------------+--------------------------------------------+-------------------------------------------+
| Name of the counties | Number of people who voted for Angela Alsobrooks | Number of people who voted for Larry Hogan | Number of people who voted for Mike Scott |
+----------------------+--------------------------------------------------+--------------------------------------------+-------------------------------------------+

**Senate By County:**

+--------------------+---------------------------------------------------------------------+----------------------------------+--------------------------------------------+----------------------------------------------+-----------------------------------------------------+----------------------------------+---------------------------------------------------+
| name               | party                                                               | early_voting                     | election_day                               | mail_in_ballot                               | provisional                                         | total                            | Percentage                                        |
+--------------------+---------------------------------------------------------------------+----------------------------------+--------------------------------------------+----------------------------------------------+-----------------------------------------------------+----------------------------------+---------------------------------------------------+
| Name of Candidates | Name of Party                                                       | Number of people who voted early | Number of people who voted on election day | Number of people who mailed in their ballots | Number of people who filled out provisional ballots | Total number of people who voted | Percentage of people who voted for each candidate |
|                    |                                                                     |                                  |                                            |                                              |                                                     |                                  |                                                   |
|                    | Democratic/ Republican/ Libertarian/ Other Candidates/ Unaffiliated |                                  |                                            |                                              |                                                     |                                  |                                                   |
+--------------------+---------------------------------------------------------------------+----------------------------------+--------------------------------------------+----------------------------------------------+-----------------------------------------------------+----------------------------------+---------------------------------------------------+

\

**Limits of the data:** None of the datasets have demographic data such as the voter's race or age, so we had to change out questions to consider that when trying to figure out what to ask. Additionally, for question 3 we have considered selecting just one other state to compare MD to rather than looking at all the flip states due to time. 

**Active Voters:** 

+---------------------+----------------------------------------------------+-----------------------------------------------------+------------------------------------------------------+------------------------------------------------+-------------------------------------------------------+---------------------------------------------------------------------------+------------------------------------------------+-------------------------------------+
| County              | DEM                                                | REP                                                 | LIB                                                  | OTH                                            | UNA                                                   | NLM                                                                       | GRN                                            | TOTAL                               |
+---------------------+----------------------------------------------------+-----------------------------------------------------+------------------------------------------------------+------------------------------------------------+-------------------------------------------------------+---------------------------------------------------------------------------+------------------------------------------------+-------------------------------------+
| Names of the county | Number of people who voted for Democrats by county | Number of people who voted for Republican by county | Number of people who voted for Libertarian by county | Number of people who voted for Other by county | Number of people who voted for Unaffiliated by county | Number of people who voted for the National Liberation Movement by county | Number of people who voted for Green by county | Total of people who voted by county |
+---------------------+----------------------------------------------------+-----------------------------------------------------+------------------------------------------------------+------------------------------------------------+-------------------------------------------------------+---------------------------------------------------------------------------+------------------------------------------------+-------------------------------------+

Q1: How many people had split ballots in Maryland?

```{r}
pres_votes <- md_pres_vice_data|>
  select(party, total)|>
  mutate(pres_votes= total)|>
  select(party, pres_votes)|>
  filter(str_detect(party, "Democratic|Republican"))

sen_votes <- senate_overall|>
  select(party, total)|>
  mutate(sen_votes= total)|>
  select(party, sen_votes)|> 
 filter(str_detect(party, "Democratic|Republican"))

sen_votes <-sen_votes |>
 group_by(party) |>
 summarise(sen_votes = sum(sen_votes, na.rm = TRUE)) 

split<- pres_votes |> 
  inner_join(sen_votes, by = "party")|>
  mutate(pct_pres = (pres_votes/sum(pres_votes)*100))|>
  mutate(pct_sen = (sen_votes/sum(sen_votes)*100))|>
  mutate(diff = pct_pres - pct_sen)
  

print(split)
```

Answer: More people voted in the presidential election than in the senate election.

Q2: How many registered voters compared to voter turnout by state and counties?

```{r}
pres_election_by_md_counties <- pres_election_by_md_counties|>
  rename(county = jurisdiction)



turnout <- pres_election_by_md_counties|>
  inner_join(voter_registration, by ="county")|>
  rename(democratic = kamala_d_harris_and_tim_walz_democratic)|>
  rename(republican = donald_j_trump_and_jd_vance_republican)|>
  rename(libertarian = chase_oliver_and_mike_ter_maat_libertarian)|>
  rename(green = jill_ellen_stein_and_rudolph_ware_green)|>
  rename(unaffiliated = robert_f_kennedy_jr_and_nicole_shanahan_unaffiliated)



turnout <- turnout|>
  mutate(total_votes = democratic+republican+libertarian+green+unaffiliated)|>
  mutate(diff= total - total_votes)

mean_value <- mean(turnout$diff, na.rm = TRUE)
```

Q3: How does the voter turnout compare in counties with different political majorities?

```{r}
turnout <- turnout|>
  mutate(
    win= case_when(
    (democratic > republican) ~ "TRUE",
    (democratic < republican) ~"FALSE"
  ))

turnout_county <- turnout |>
  select(county, diff, win)|>
  arrange(diff)
```

Q4: Based on current voting data, how might voting patterns in Maryland change in upcoming elections?

```{r}
elect_2008 <- elect_2008|>
  select(party, percentage)|>
  mutate(percentage= as.numeric(parse_number(percentage)))

elect_2012 <- elect_2012|>
  select(party, percentage)|>
  mutate(percentage= as.numeric(parse_number(percentage)))

elect_2016 <- elect_2016|>
  select(party, percentage)|>
  mutate(percentage= as.numeric(parse_number(percentage)))

elect_2020 <- elect_2020|>
  select(party, percentage)|>
  mutate(percentage= as.numeric(parse_number(percentage)))

elect_2024 <- md_pres_vice_data|>
  select(party, percentage)|>
  mutate(percentage= as.numeric(parse_number(percentage)))


patterns <- elect_2008 |>
  inner_join(elect_2012, by="party")|>
  rename(elect_2008 = percentage.x)|>
  rename(elect_2012 = percentage.y)|>
  inner_join(elect_2016, by="party")|>
  rename(elect_2016 = percentage)|>
  inner_join(elect_2020, by="party")|>
  rename(elect_2020 = percentage)|>
  inner_join(elect_2024, by="party")|>
  rename(elect_2024 = percentage)

patterns <- patterns|>
  mutate(prediction = (elect_2008+elect_2012+elect_2016+elect_2020+elect_2024)/5)
```

Q5: What senate election had the highest turnout by county?

```{r}

```
