---
title: "Baltimore Homicides"
author: "Peter, Liam, Andrew"
date: "10/17/24"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, [Baltimore homicides](https://www.washingtonpost.com/investigations/as-police-struggle-to-solve-homicides-baltimore-residents-see-an-open-season-for-killing/2018/12/26/7ee561e4-fb24-11e8-8c9a-860ce2a8148f_story.html)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
homicide_data <- read.csv('data/homicide-data.csv')


```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: “City police made an arrest in 41 percent of homicides in 2014; last year, the rate was just 27 percent, a 14 percentage point drop.”
-   **Analysis summary**: I think we were able to get close to the finding but could not confirm it 100%. We were able to filter by year but the last step in terms of percent drop I was not able to figure out.

```{r}
# Put code to reverse engineer sentence here
homicide_data <- homicide_data|>
  rename(date = reported_date)|>
  mutate(date=ymd(date))

# Display results of code below this codeblock

```

```{r}
baltimore_homicides <- homicide_data |>
  filter(str_detect("Baltimore", city)) |>
  mutate(year= lubridate:: year(date))
```

```{r}
baltimore_homicides <- baltimore_homicides |>
  mutate(
    arrest_made = case_when(
      str_detect(disposition, "Open/No arrest") ~ "NO",
      str_detect(disposition, "Closed without arrest") ~ "NO",
      str_detect(disposition, "Closed by arrest") ~ "YES"
    )
  )
```

This was as far as I could get but I could not get it to be one table and I could not figure out the percentage change. I thought about converting the "YES" and "NO" to 1s and 2s and change if from chr to num but I didn't think that would work as I think I would just get back a bunch of 1s and 2s without the read percent change.

```{r}
baltimore_homicides|>
  filter(str_detect(year, "2014"))

baltimore_homicides|>
  filter(str_detect(year, "2017"))
```

### Sentence 2

-   **Sentence text**: Of 50 of the nation’s largest cities, Baltimore is one of 34 where police now make homicide arrests less often than in 2014, according to a Washington Post analysis.
-   **Analysis summary**: We were able to show that Baltimore was one of the 34 cities where police now make homicide arrests less often than in 2014.

```{r}

homicide_arrests_2014 <- homicide_data |> filter (str_detect(year(date), "2014"))

homicide_arrests_2014 <- homicide_arrests_2014 |> filter(str_detect(disposition,"Closed by arrest")) |> group_by (city) |> summarise(arrests2014 = sum(n())) |> arrange(desc(arrests2014))

homicides_2014 <- homicide_data |> filter(str_detect(year(date), "2014")) |> group_by (city)|> summarise (homicides2014 = sum(n()))

homicides_and_arrests_2014 <- homicides_2014 |> full_join (homicide_arrests_2014, join_by(city))

homicides_and_arrests_2014 <- homicides_and_arrests_2014 |> mutate(
  arrest_rate_2014 = arrest_rate_2014 <- (arrests2014/homicides2014) * 100
)

homicide_arrests_2017 <- homicide_data |> filter(str_detect(year(date), "2017"))

homicide_arrests_2017 <- homicide_arrests_2017 |> filter(str_detect(disposition,"Closed by arrest")) |> group_by (city) |> summarise(arrests2017 = sum(n())) |> arrange(desc(arrests2017))

homicides_2017 <- homicide_data |> filter(str_detect(year(date), "2017")) |> group_by (city) |> summarise (homicides2017 = sum(n()))

homicides_and_arrests_2017 <- homicides_2017 |> full_join (homicide_arrests_2017, join_by(city))

homicides_and_arrests_2017 <- homicides_and_arrests_2017 |> mutate(
  arrest_rate_2017 = arrest_rate_2017 <- (arrests2017/homicides2017) *100)

combined_homicide_arrest_rates <- homicides_and_arrests_2014 |> inner_join(homicides_and_arrests_2017, join_by(city))

combined_homicide_arrest_rates <- combined_homicide_arrest_rates |> mutate(
  arrests_less_often = case_when(
    arrest_rate_2014 >= arrest_rate_2017 ~ "YES",
    arrest_rate_2014 < arrest_rate_2017 ~ "NO"
  )
)

combined_homicide_arrest_rates |> filter(str_detect(arrests_less_often, "YES")) |> summarise (cities_with_fewer_arrests = sum(n()))
```

### Sentence 3

-   **Sentence text**: [There is an “indisputable” Freddie Gray effect] “And while most cities saw their arrest rates drop gradually, Baltimore’s decline was sudden — plummeting 15 percentage points in 2015, after Gray’s death, the largest single-year drop for any city already solving less than half its homicides
-   **Analysis summary**: We want to show that from 2014 to 2015, Baltimore's arrest rate dropped by 15%. Which we were able to do with following code.

```{r}
b2014_arrest <- baltimore_homicides |>
  filter(str_detect(year(date), "2014"))|>
  filter(str_detect(disposition, "Closed by arrest"))|>
  summarise(n())

b2014_homicides <- baltimore_homicides |>
  filter(str_detect(year(date), "2014"))|>
  summarise(n())

b2015_arrest <- baltimore_homicides |>
  filter(str_detect(year(date), "2015"))|>
  filter(str_detect(disposition, "Closed by arrest"))|>
  summarise(n())

b2015_homicides <- baltimore_homicides |>
  filter(str_detect(year(date), "2015"))|>
  summarise(n())

b2014_arrest/b2014_homicides
b2015_arrest/b2015_homicides


```

### Sentence 4

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
