---
title: "lab_12"
author: "Mitchell Hang"
date: "2023-05-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   tidytext and our usual libraries

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidytext)
library(tidyverse)
library(dplyr)
library(janitor)
library(lubridate)
library(rvest)
```

## Questions

##### **Q1.** You've been assigned to report a story about the leading reasons that Maryland attorneys get sanctioned by the state for misconduct. The state [publishes lists of sanctions](https://www.courts.state.md.us/attygrievance/sanctions) that contain a short text description about the situation. Load the CSV file in the data folder containing records from fiscal year 2011 onwards. Make a list of unique words from the text column, then following the example in the pre_lab, remove common "stop words" from that list and create a list of the top 10 words containing the percentage of occurrences each word represents. What's the leading word in that answer and, broadly, what do you think the top 10 words describe?

```{r}
#Load the data
md_attorney_sanctions <- read_csv("data/md_attorney_sanctions.csv")
```

```{r}
a_list_of_words <- c("the", "and", "is", "this", "that", "for", "to", "by")
unique(a_list_of_words)
```

```{r}
data("stop_words")

stop_words <- stop_words |> 
  add_row(word = "client") |> 
  add_row(word = "justice") |> 
  add_row(word = "respondent") |>
  add_row(word = "clients")|>
  add_row(word = "attorney")|>
  add_row(word = "representation")

```

```{r}
word_counts <- md_attorney_sanctions|>
  unnest_tokens(word, text)|>
  anti_join(stop_words)|>
  count(word, sort = TRUE)|>
  mutate(percentage = n / sum(n) * 100)

top_10_words <- word_counts |>
  slice_max(n, n = 10)

top_10_words
```

**A1. The leading word is failing, I think that the words together are describing how someone failed to do something. whether it be mistrust, misconduct, etc.**

------------------------------------------------------------------------

##### **Q2.** Let's move beyond single words to phrases. Make a list of the top 10 three-word phrases, called trigrams, based on the example from the pre_lab (you'll need to modify the example code to do this). What's the top trigram and how often does it appear? What does that phrase mean in legal terms?

```{r}
# Check the task 12 on Pre-lab 11 to see if you can use that code
trigram_counts <- md_attorney_sanctions |>
  unnest_tokens(trigram, text, token = "ngrams", n = 3) |>
  count(trigram, sort = TRUE)|>
  mutate(percentage = n / sum(n) * 100)

top_10_trigrams <- trigram_counts|>
  slice_max(n, n = 10)

top_10_trigrams
```

**A2. My top two were "for failing to" and "by consent on" meaning having not done something you were supposed to or with approval on something**

------------------------------------------------------------------------

##### **Q3.** Let's drop back down to more traditional text analysis - take the top trigram from Q2 and write code to see how many times it occurs in the text column in each fiscal year. What do you think the answer produced by your code suggests? What else could you do to try and clarify the most important reasons attorneys get sanctioned?

```{r}
phrase_counts <- md_attorney_sanctions|>
  filter(str_detect(text, "for failing to"))|>
  group_by(fiscal_year)|>
  summarize(count = n())

phrase_counts
```

**A3. it really spiked since 2011 but then evened out to around the high 40s by 2020. I tried to filter out words like to, and for but for some reason it did not carry over. I think that could have yielded different results but failure to do something also does not surprise me for being at the top of the list.**
