---
title: "lab_01"
author: "Derek Willis"
adapted by: "Daniel Trielli"
date: "2024-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your class reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the UMD courses data.

**Task** Run the codeblock below to load the data, which is contained in an RDS file.

```{r}
umd_courses <- read_rds("lab_01.rds")
```

## Answer questions

### **Q1** Which department has the highest mean (average) number of seats for its classes? You need to show the department with the highest average first. Which one is the most surprising/interesting to you?

```{r}
umd_courses |>
  group_by(department) |>
  summarise(
    mean_seats = mean(seats)
  ) |>
  arrange(desc(mean_seats))
```

**A1**

Engineering, Cooperative Education has the highest mean (306.666...) seats. I was surprised that Air Science was higher up than courses like Computer Science or Mathematics.

------------------------------------------------------------------------

### **Q2** Find out the greatest number of times any individual class has been offered in our dataset. You must show the greatest number of classes first.

```{r}
umd_courses |>
  group_by(title) |>
  summarize(
    count_classes = n(),
    sum_term = sum(term)
  ) |>
  arrange(desc(count_classes))
```

**A2**

Doctoral Dissertation Research has been offered the most (1,303).

------------------------------------------------------------------------

### **Q3** Count the number of courses for each instructor and department. Which instructor has taught the most?

```{r}
umd_courses |>
  group_by(instructors) |>
  summarize(
    count_courses = n()
  ) |>
  arrange(desc(count_courses))
```

**A3**

With the exception of courses that have not that have not been assigned an instructor (764), or those labeled 'N/A,' (10,996) Elsa Barkley Brown has taught the most about of courses (73).

------------------------------------------------------------------------

### **Q4** Given the results for Q3, what would have made it easier to find the answer? You don't have to know *how* to accomplish this; I want you to describe in words what would make the answer easier to see.

```{r}

```

**A4**

I'm sure there is a way, but I would have sorted them the same way except edit the code so it does not count courses that have not been assigned someone. So the ones labeled "TBA' or 'N/A' would not be part of the count.

------------------------------------------------------------------------

### **Q5** Take one of the questions you listed in the final question of the pre_lab and try to answer it here in code. If you don't know how to fully answer it, write what code you can and describe in words what else you would need to do. If you got an answer, is it interesting?

```{r}
umd_courses |>
  group_by(title) |>
  summarize(
    count_classes = n(),
    sum_term = sum(term)
  ) |>
  arrange(desc(count_classes))
```

**A5**

For my third question in the pre_lab, I was wondering how many of the courses were possibly work.independent study. For this I could use the same code as question 2. Again the only thing I would change is finding a way to sort by specific class. Like sort by independent study or sort by internship. But I am not sure if you can do this or if you have to sort by title and then scroll till you find the title you are looking for.

The answer to this is that Independent study has 443 and internships have 79.
